---
title: "Preliminary Analysis of RTC Data"
author: "Matthew Knowles"
date: "`r Sys.Date()`"
output:
    pdf_document: default
html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(purrr)
set.seed(5)
accidents <- read.csv("../Data/dft-road-casualty-statistics-accident-2020.csv")
vehicles <- read.csv("../Data/dft-road-casualty-statistics-casualty-2020.csv")
casualties <- read.csv("../Data/dft-road-casualty-statistics-vehicle-2020.csv")
```

# 1: Data Cleaning

We begin by converting to a binary variable by severity. Accidents with a severity of 1 stay as such, and others are converted to a 0. 

```{r}
accidents <- accidents %>% 
    mutate(bin_severity = ifelse(.data$accident_severity == 1, 1, 0)) %>% 
    filter(light_conditions != -1)
```

We now reduce the size of the dataset to contain only variables we care about

```{r}
accidents <- accidents %>% 
    select(
        "bin_severity",
        "first_road_class",
        "day_of_week",
        "road_type",
        "speed_limit",
        "junction_detail",
        "second_road_class",
        "light_conditions",
        "weather_conditions",
        "road_surface_conditions"
    )
```

We want all variables, except speed limit to be factors.

```{r}
accidents <- accidents %>%
    mutate(across(where(is.integer), as.factor))
accidents$speed_limit <- as.integer(accidents$speed_limit)

head(tibble(accidents))
```


We can now also create a train and test set for later on.

```{r}
accidents$id <- 1:nrow(accidents)
accidents_train <- accidents %>% sample_frac(0.7)
accidents_test <- anti_join(accidents, accidents_train, by = "id")
```

# 2: Fitting

```{r}
fit <- glm(
    data = accidents_train,
    formula = bin_severity ~ .,
    family = binomial(link = "logit")
)
```

Let us take a look at the summary of this model.

```{r}
summary(fit)
```

```{r}
hist(fit$fitted.values)
```

# 3: Predicting Unseen Values

```{r}
predict_fit <- predict(
    fit,
    newdata = accidents_test,
    type = "response"
)
```

```{r}
hist(predict_fit)
```
